Minimal trap spaces of Logical models are maximal siphons of their Petri net encoding

### 2022-06-16 cmsb22.tex
Comments on the benchmark conditions and back =15p

To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}.
Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet). We ran all the benchmarks on an apple laptop whose environment is CPU:\@ Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1. Finally, we set a time limit of two minutes for each model.

@@

Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
\revise{Since the ASP code generated is also quite declarative by nature, it is also possible to add to it supplementary constraints coming from the modeller in case one is looking for specific attractors.
Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.}

@@

Though we benchmarked this new approach against the state-of-the-art tool PyBoolNet, there are many more evaluations that we plan to do in the future.
First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.

@@

In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}. Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future. They could replace our ASP program if they outperform it.

\revise{However, the current method appears to already perform very well even on the biggest models we have considered.}
We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.

>>>>

To solve the ASP problems, we used the same ASP solver CLINGO~\cite{DBLP:journals/aicom/GebserKKOSS11} and the same configuration as that used in PyBoolNet~\cite{klarner2015computing,klarner2017pyboolnet}.
Specifically, we used the configuration \texttt{-heuristic=Domain -enum-mod=domRec -dom-mod=3} (subset maximality, equivalent to the deprecated \texttt{--dom-pref=32 --heuristic=domain --dom-mod=7} used by PyBoolNet). We ran all the benchmarks on an apple laptop whose environment is CPU:\@ Intel\textregistered\ Core\texttrademark\ i7 1.20GHz x 4, 16 GB DDR4 RAM, MacOS 12.3.1. Finally, we set a time limit of two minutes for each model.
\revise{Note that we did not get the opportunity to run this benchmark on a proper dedicated computing workstation, the results are therefore only indicative of global trends and should not be interpreted as precise performance.
This is also why in some rare cases finding all minimal trap spaces was faster than with the limit to 1000.}

@@

Note that even with a very restricted time-limit of two minutes, it was possible with the proposed technique to find \emph{all} minimal trap spaces of small models (roughly under 130 nodes, i.e., considered as quite big up to now).
Though it might seem impractical to handle tens of thousands of such possible complex attractors in a manual way, i.e., to compare them to specific experimental conditions and corresponding data, we hope that an automatic analysis of such attractors might become possible with systematic verification methods, not unlike that described in~\cite{hernandez2020computational}.
\revise{Since the ASP code is declarative by nature, it is also possible to add to it supplementary constraints coming from the modeler in case one is looking for specific attractors.
Finally, sampling from the ASP-generated solutions as is done in~\cite{chevalier2020synthesis} would allow for a different type of exploration.}

@@

Though we benchmarked this new approach against state-of-the-art \revise{tools}, there are many more evaluations that we plan to do in the future.
First, the BioLQM platform that we are using is providing another implicant-based method using BDDs in \url{http://colomoto.org/biolqm/doc/tools-trapspace.html} and though we expect it to behave mostly like PyBoolNet, that remains to be checked.
Note that this also raises the question of replacing altogether the BioLQM preprocessing step we use to obtain the Petri net encoding, since its use of BDDs might not be optimal for that step.
The trade-off between a small Petri net, and hence small ASP, and the time it takes to compute it (in other words, the trade-off of allowing redundant constraints) has to be evaluated in depth.

@@

In addition, there are possibly other methods for computing maximal conflict-free siphons in Petri nets, like SAT/MaxSAT approaches~\cite{nabli2016enumerating}. Although these approaches do not directly support the maximal conflict-free siphon computation now, we plan to investigate them in the future. They could replace our ASP program if they outperform it.
\revise{However, the current method appears to already perform very well even on the biggest models we have considered.}
% We thus plan to propose its inclusion in the CoLoMoTo effort to make it readily available to modellers.
